---
title: "HW#_Q4"
author: "Daniil Deych"
date: "`r Sys.Date()`"
output: pdf_document
          keem_md: TRUE
---
# Question 4

```{r, echo=FALSE}
install.packages('mapdata')
install.packages('sf')

# Install and load necessary packages
install.packages(c("ggplot2", "maps", "ggmap", "maptools", "rgdal"))
library(ggplot2)
library(maps)
library(ggmap)
library(maptools)
library(rgdal)


library(tidyverse)
library(rpart)
library(rpart.plot)
library(rsample) 
library(dplyr)

library(randomForest)
library(lubridate)
library(modelr)

library(gbm)
library(pdp)
library(MASS)

library(readr)
library(ggplot2)
library(reshape2)
library(maps)
library(mapdata)


CAhousing <- read_csv("CAhousing.csv")
```
To find our best predictive model, we elected to use three different methods. Since our total number of features isn't that high - only 9 after standardizing by household. For that reason some version of a tree model is likely to work best.

To test that approach we also included a linear model to compare against.

The first model we attempt is a Random Forest Model, which we then compare to the General Boosted Regression approach.

## Random Forest Model

```{r, echo=FALSE}
#Creating variables demonstrating averages of statistics per household
colnames(CAhousing)[c(1, 2)] <- c("long", "lat")

CAH <- CAhousing %>%
  mutate(avg_rooms = totalRooms/households, 
         avg_bedrooms = totalBedrooms/households,
         avg_house_pop = population/households) 

#split data
CAH_split = initial_split(CAH)
CAH_train = training(CAH_split)
CAH_test = testing(CAH_split)

CAH_forest = randomForest(medianHouseValue ~ housingMedianAge + medianIncome + 
                          avg_rooms + avg_bedrooms + avg_house_pop + long
                          + lat, data = CAH_train,
                          importance = TRUE)

# shows out-of-bag MSE as a function of the number of trees used
plot(CAH_forest)

#Forest Prediction for price given location
partialPlot(CAH_forest, as.data.frame(CAH_test), long, las=1)
partialPlot(CAH_forest, as.data.frame(CAH_test), lat, las=1)
```

## General Boosted Regression Model

We make a build set and a check set to adjust parameters.

```{r, echo=FALSE, message=FALSE}
CAH_train_split = initial_split(CAH_train)
CAH_boost_build = training(CAH_train_split)
CAH_boost_check = testing(CAH_train_split)


CAH_boost1 = gbm(medianHouseValue ~ housingMedianAge + medianIncome +
                     avg_rooms + avg_bedrooms + avg_house_pop + long
                     + lat, data = CAH_boost_build,
              interaction.depth=11, n.trees=1000, shrinkage=.08)
 
CAH_boost2 = gbm(medianHouseValue ~ housingMedianAge + medianIncome +
                    avg_rooms + avg_bedrooms + avg_house_pop + long
                  + lat, data = CAH_boost_build,
                     interaction.depth=14, n.trees=1000, shrinkage=.08)

CAH_boost3 = gbm(medianHouseValue ~ housingMedianAge + medianIncome +
                   avg_rooms + avg_bedrooms + avg_house_pop + long
                 + lat, data = CAH_train,
                    interaction.depth=12, n.trees=1000, shrinkage=.08)

rmse(CAH_boost1, CAH_boost_check)
rmse(CAH_boost2, CAH_boost_check)
rmse(CAH_boost3, CAH_boost_check)
```
After adjusting the parameters, we conclude that the interaction depth of 12 is most optimal.

# Plot of predictions for y given location 

```{r, echo=FALSE, fig.show=FALSE }
p1 = pdp::partial(CAH_boost3, pred.var = 'long', n.trees=1000)
ggplot(p1) + geom_point(mapping=aes(x=long, y=yhat))

p2 = pdp::partial(CAH_boost3, pred.var = 'lat', n.trees=1000)
ggplot(p2) + geom_point(mapping=aes(x=lat, y=yhat))
```
As a comparison, we also added a linear model.

# Linear model

```{r, echo=FALSE}
CAH_lm = lm(data = CAH_train, medianHouseValue ~ housingMedianAge + medianIncome +
              avg_rooms + avg_bedrooms + avg_house_pop + long
            + lat )
CAH_stepwise = stepAIC(CAH_lm, direction = "both", 
                        trace = FALSE)
summary(CAH_stepwise)
```
Finally we compared the out-of-sample error for the 3 approaches against our test set.

# RMSE test

```{r, echo=FALSE}
rmse(CAH_boost3, CAH_test)
rmse(CAH_forest, CAH_test)
rmse(CAH_lm, CAH_test)
```
After comparing RMSE values between GBM, RandomForest and linear models, the lowest RMSE value belongs to our GBM model. 

GBM That will act as our best predictive model, which provided the following results.

# Plot of the Original Data
```{r, echo=FALSE}
ca_data <- map_data("state", region = "california")

california_map <- ggplot(ca_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "grey50", color = "white") +
  theme_void()

colnames(CAH)[c(1, 2)] <- c("long", "lat") 

CAH_MedianHouseValue = CAH[c(1,2,9)]

california_MedianHouseValue <- ggplot() +
  geom_polygon(data = ca_data, aes(x = long, y = lat, group = group), fill = "white", color = "black") +
  geom_tile(data = CAH_MedianHouseValue, aes(x = long, y = lat, fill = medianHouseValue), alpha = 1) +
  scale_fill_gradient(low = "blue", high = "red") +
  coord_equal() +
  labs(title = "California Median House Value Map", fill = "Median House Value")

california_MedianHouseValue
```
# Plot of our best Prediction Model
```{r, echo=FALSE}
colnames(CAH_test)[c(1, 2)] <- c("long", "lat")

CAH_MedianHouseValue_test = CAH_test[c(1,2,9)] %>%
  mutate(ln_medianHouseValue = log(medianHouseValue))

california_MedianHouseValue_GBM <- ggplot() +
  geom_polygon(data = ca_data, aes(x = long, y = lat, group = group), fill = "white", color = "black") +
  geom_tile(data = CAH_MedianHouseValue_test, aes(x = long, y = lat, color = medianHouseValue), alpha = 1) +
  scale_fill_gradient() +
  coord_equal() +
  labs(title = "Predictive California Median House Value Map", fill = "Median House Value")

california_MedianHouseValue_GBM
```
# Error/residuals Plot
```{r, echo=FALSE, warning=FALSE}

CAH_MedianHouseValue_test = CAH_MedianHouseValue_test %>%
  mutate(value_predgbm = predict(CAH_boost3, CAH_test, n.trees = 1000),pred_errorgbm=(medianHouseValue-value_predgbm)^2) 

CAH_MedianHouseValue_test = CAH_MedianHouseValue_test %>%
  mutate(sqpred_errorgbm=sqrt(pred_errorgbm))

california_MedianHouseValue_ErrorGBM <- ggplot() +
  geom_polygon(data = ca_data, aes(x = long, y = lat, group = group), fill = "white", color = "black") +
  geom_tile(data = CAH_MedianHouseValue_test, aes(x = long, y = lat, color = sqpred_errorgbm), alpha = 1) +
  scale_fill_gradient(low = "lightblue", high = "red") +
  coord_equal() +
  labs(title = "Error/Residual Map")

california_MedianHouseValue_ErrorGBM

```



